---
title: "ggplot2 Cheatsheet"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
author: "Jeff Horowitz"
output:
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background
This is a working guide to create plots using the ggplot2 library. It uses the capital bikeshare data which is publically available here: https://capitalbikeshare.com/system-data. 

# Load Libraries

Load in all libraries required for plotting.

```{r}
library(tidyverse)
library(dplyr) 
library(ggplot2)
library(ggthemes)
library(knitr)
library(DataExplorer)
library(scales)
library(extrafont)
loadfonts(device = "win")
```

# Data Loading and Pre-Processing

Begin by loading data and pre-processing, including filling in any missing values.

```{r}
bikeshare <- read_csv("bikesharedailydata.csv")

kable(head(bikeshare,5))

str(bikeshare)

bikeshare$yr <- factor(bikeshare$yr, levels = c(0, 1), labels = c("2011", "2012"))

bikeshare$season <- factor(bikeshare$season, levels = c(1,2,3,4), labels = c("Winter", "Spring", "Summer", "Fall"))

bikeshare <- rename(bikeshare, humidity = hum, month=mnth)
names(bikeshare)

colSums(is.na(bikeshare))

colnames(bikeshare)[colSums(is.na(bikeshare)) > 0] 

biketidy <- bikeshare[complete.cases(bikeshare),]
nrow(biketidy)
nrow(bikeshare)-nrow(biketidy)  

biketrack <- bikeshare[!complete.cases(bikeshare),] 
biketrack

bikeshare$season[7] <- "Winter"
bikeshare$month[10] <-1
bikeshare$season[7]
bikeshare$month[10]

plot_missing(bikeshare) 
```

# Build Theme

```{r}
mytheme <- theme(panel.background = element_blank(),
                 panel.grid = element_blank(),
                 axis.line = element_line(color="#999999"),
                 axis.ticks.y = element_blank(),
                 axis.ticks.x = element_blank(),
                 text = element_text(family = "Tahoma", size = 11),
                 plot.title = element_text(size=11))
```


# 1. Bar Chart

```{r}

ggplot(data = bikeshare, aes(x=season, y=cnt)) + 
  geom_bar(stat = "identity", color = "#1594e8", fill = "#1594e8") +
  scale_y_continuous(labels = comma) + 
  labs(title = "Bike Rentals per Season (2011 - 2012)",
       caption = "Jeff Horowitz | Data Source: Capital Bikeshare") +
  mytheme +
  theme(
    axis.title = element_blank()
  )

```

# 2. Histogram

```{r}

hist_data <- subset(bikeshare, bikeshare$yr == 2011)

hist_data_binned <- hist_data %>%
  mutate(bin = cut(cnt, breaks = 30)) %>%
  group_by(bin) %>%
  summarize(count = n()) %>%
  ungroup()

max_bin <- hist_data_binned %>%
  filter(count == max(count)) %>%
  pull(bin)

ggplot(data = hist_data, aes(x = cnt)) +
  geom_histogram(aes(fill = cut(cnt, breaks = 30) == max_bin), color = "#999999", bins = 30) +
  scale_fill_manual(values = c("FALSE" = "#1594e8", "TRUE" = "#ea6b17"), guide = "none") +
  scale_x_continuous(labels = comma) +
  labs(title = "Frequency of Bike Rentals (2011)",
       x = "Rentals",
       caption = "Jeff Horowitz | Data Source: Capital Bikeshare") +
  mytheme +
  theme(axis.title.y = element_blank())

```

# 3. Scatterplot

```{r}

ggplot(data = bikeshare, aes(x=cnt, y=temp)) + 
  geom_point(color = "#1594e8") +
  scale_x_continuous(labels = comma) + 
  labs(title = "Relationship of Temperature and Bike Rentals (2011 - 2012)",
       x = "Rentals",
       y = "Temperature (F)",
       caption = "Jeff Horowitz | Data Source: Capital Bikeshare") +
  mytheme
```

# 4. Line Chart

```{r}

line_data <- subset(bikeshare, bikeshare$season == "Summer" & bikeshare$yr == 2011)

ggplot(data = line_data, aes(x=1:nrow(line_data), y=cnt)) + 
  geom_line(color = "#1594e8") + 
  scale_y_continuous(labels = comma) + 
  labs(title = "Bike Rentals per Day in Summer (2011)",
       x = "Days",
       y = "Number of Rentals",
       caption = "Jeff Horowitz | Data Source: Capital Bikeshare") + 
  mytheme
```

# 5. Boxplot

```{r}

box_data <- subset(bikeshare, bikeshare$yr == 2011)
box_data$month <- as.factor(box_data$month)
levels(box_data$month) <- c("Jan","Feb","March","April","May","June","July","Aug","Sept","Oct","Nov","Dec")

ggplot(data = box_data, aes(x=month, y=cnt)) +
  geom_boxplot(color = "#1594e8") +
  scale_y_continuous(labels = comma) +
  labs(title = "Bike Rentals per Month (2011)",
       caption = "Jeff Horowitz | Data Source: Capital Bikeshare") + 
  mytheme +
  theme(axis.title = element_blank())
```

# 6. Density Plot

```{r}

ggplot(data = bikeshare, aes(x=cnt)) +
  geom_density(color = "#1594e8", fill = "#1594e8") + 
  scale_x_continuous(labels = comma) + 
  labs(title = "Density Plot of Bike Rentals (2011 - 2012)",
       x = "Rentals",
       y = "Density",
       caption = "Jeff Horowitz | Data Source: Capital Bikeshare") +
  geom_vline(xintercept = round(mean(bikeshare$cnt),2), size = 1, color = "#15e82e", linetype = "dotted") +
  mytheme
  
```

# 7. Stacked Area

```{r}
seasonal_data <- subset(bikeshare, bikeshare$yr == 2011)
seasonal_data <- seasonal_data %>%
  group_by(month) %>%
  summarise(casual = sum(casual, na.rm = TRUE), 
            registered = sum(registered, na.rm = TRUE)) %>%
  pivot_longer(cols = c(casual, registered), names_to = "user_type", values_to = "count")

seasonal_data$month <- as.factor(seasonal_data$month)
levels(seasonal_data$month) <- c("Jan","Feb","March","April","May","June","July","Aug","Sept","Oct","Nov","Dec")
seasonal_data$user_type <- as.factor(seasonal_data$user_type)

ggplot(data = seasonal_data, aes(x = month, y = count, group = user_type, fill = user_type)) +
  geom_area(position = "stack", color = "#999999") +
  scale_fill_manual(values = c("casual" = "#ea6b17", "registered" = "#1594e8")) +
  scale_y_continuous(labels = comma) +
  labs(title = "Bike Rentals per Month (2011)", 
       caption = "Jeff Horowitz | Data Source: Capital Bikeshare",
       fill = "Type") +
  mytheme +
  theme(axis.title = element_blank(),
        legend.position = c(.12,.85),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8))
```

